Physics of Equilibrium

### 2019-05-10 distill_draft_notes/Draft/Draft.tex
changes latex file

The Boltzmann distribution is known from physics.
and it describes the statistics of physical configurations of a system that is in an equilibrium at some temperature. We do not explicitely write the temperature because we either set it to $T=1$ or it is already absorbed in the parameters $\lambda_k$ of the model. But it is important that we are not looking at systems with zero temperature because in this case a physical system always will equilibrate to its state of minimum energy. Allowing a non zero temperature automatically leads to states that are a thermal distribution which means we would observe many possible configurations that are not only of minimum energy. In the extreme case of very high temperature all the configurations even get equally likely.

@@

\subsubsection{Paragraph for Boltzman distribution at zero temperature}

@@

So far this result is general and can be applied to any constraint, the choice of $\langle r_k \rangle_x$ will in the end define the physics of the model. The most commonly used physical model in machine learning that defines the underlying behaviour of the Boltzmann distribution is a so called Ising model, which is defined through a energy function $E(\vec {\sigma}) = \sum_{i,j} w_{i,j} \sigma_i \sigma_j + \sum_i h_i \sigma_i$. 
The sufficient statistic of the Boltzmann distribution of an Ising model are the single-spin expectation values $\langle \sigma_i \rangle$ and the two-spin correlations $\langle \sigma_i \sigma_j \rangle$. This means that these two constraints suffice to fully determine the parameters of the classical Ising energy function from the data.
Generally for the exponential family a statistic $T(x)$ is sufficient, if we can write the probability $p(x)$ as

@@

@@

@@

I physics low energies are preferred over high energies. 
And therefore if we initialize a system randomly it will eventually equilibrate to a low energy configuration or if we are at $T=0$ to the minimum energy.
If we imagine to have a spin model implemented as a physical device and we can initialise the spins in a random way, this spin configuration would automatically go to a configuration of lower energy.
Therefore in a real physical system, sampling from a system is only a matter of initialising it and waiting until it is equilibrate.

@@

In ML we don't have an equilibration process that drives our system into a low energy configuration. But there are several methods how we can imitate this behaviour.
To sample from a Hopfield network one takes a random input vector $\vec{v}$ and passes it through the network several times and updates each node after another and then determines the update of a node via the update rule $v_i = 1,~ if~\sum_j W_{ij} v_j \geq 0$. The outcome in this case is deterministic and one can stop this process after two iterations through the nodes did not lead to any change.

>>>>

The Boltzmann distribution is known from physics and it describes the statistics of physical configurations of a system that is in an equilibrium at some temperature in this case the probability of a configuration $\vec{x}$ is given as $p(\vec{x}) = \frac{1}{Z} \exp{- \frac{1}{T} E(\vec{x})}$. Where $E(\vec{x})$ is a function that connects the configuration $\vec{x}$ with the real line, which is also called the energy of the system. Before we did not explicitely write the temperature because we either set it to $T=1$ or it is already absorbed in the parameters $\lambda_k$ of the model. But it is important that we are looking at systems with non-zero temperature because for zero temperature a physical system will equilibrate to its state of minimum energy. Allowing a non zero temperature automatically leads to states that are a thermal distribution which means we would observe many possible configurations that are not only of minimum energy. In the extreme case of very high temperature all the configurations even get equally likely.

@@

@@

So far this result is general and can be applied to any constraint, the choice of $\langle r_k \rangle_x$ will in the end define the physics of the model. The most commonly used physical model in machine learning that defines the underlying behaviour of the Boltzmann distribution is a so called Ising model, which is defined through a energy function $E(\vec {\sigma}) = \sum_{i,j} w_{i,j} \sigma_i \sigma_j + \sum_i h_i \sigma_i$. 
The sufficient statistic of the Boltzmann distribution of an Ising model are the single-spin expectation values $\langle \sigma_i \rangle$ and the two-spin correlations $\langle \sigma_i \sigma_j \rangle$. This means that these two constraints suffice to fully determine the parameters of the classical Ising energy function from the data.
Generally for the exponential family, a statistic $T(x)$ is sufficient, if we can write the probability $p(x)$ as

@@

\subsubsection{Paragraph for Boltzman distribution at zero temperature??}

@@

\subsection{Sampling}

\textbf{We should add the plot here with the nodes that equilibrate, to show that equilibration makes random inputs equilibrate. The problem here is, that we have not yet introduced the BM. We are still with general spin models.}

@@

I physics low energies are preferred over high energies. 
And therefore if we initialize a system randomly it will eventually equilibrate to a low energy configuration or if we are at $T=0$ to a local minimum of the energy.
If we imagine to have a spin model implemented as a physical device and we can initialise the spins in a random way, this spin configuration would automatically go to a configuration of lower energy.
Therefore in a real physical system, sampling from a system is only a matter of initialising it and waiting until it is equilibrate.

@@

In ML we don't have an equilibration process that drives our system into a low energy configuration. But there are several methods how we can imitate this behaviour.
To sample from a Hopfield network one takes a random input vector $\vec{v}$ and updates each node after another and then determines the update of a node via the update rule $x_i = 1,~ if~\sum_j W_{ij} x_j \geq h_i$. The outcome in this case is deterministic and one can stop this process after the nodes have been iterated through two times without any update.
For non-zero temperature models, the updates work similarly, just that the update is not deterministic, but there is a probability assigned to every node $x_i$ according to its input strength $\sum_j W_{ij} x_j + h_i$. One can imagine that
